Accelerating parameter inference with graphics processing units

### 2019-02-02 paper/paper.tex
paper.tex: finish updating BNS sec

\begin{figure*}
\caption{\label{fig:BNS:MultiIterate}\textbf{Convergence of BNS analysis}: Results for marginal posterior distributions
  of our fiducial synthetic neutron star.  Solid contours show credible intervals; solid one-dimensional distributions
  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\lambda$ and
  value of the underlying marginalized likelihood evaluations.   The corresponding dotted curves show an analysis using
  only the $m=\pm 2$ modes \editremark{perform both} \emph{Left panel } Posterior distribution
  over  $\mc$ and
  $\delta=(m_1-m_2)/M$.    \emph{Center panel}: Marginal 1d CDFs of $\mc$, showing convergence.
\emph{Right panel}: Marginal 1d empirical CDF of $\ln{\cal L}_{\rm marg}$, showing that at late times the proposed
candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$.
}
\end{figure*}

% Code configuration
In this example, we use $100$ initial points spread over the two mass dimensions, adding $20$ evaluations per
iteration.  With one instance analyzing each 20 evaluations,  we use five cores in the first iteration, followed by one
core for remaining iterations.  We choose this simple configuration to facilitate visualization.    The left and center panels of Figure  \ref{fig:BNS:MultiIterate} show posterior disributions derived
from the first several RIFT iterations.

* show convergence diagnostic of $\ln {\cal L}$ range over the data: can see few outliers.  Easy way to assess agreement
and stability

** can also do L2 difference between likelihoods estimated with different data

* COST OF OVERALL ANALYSIS, AVERAGED OVER WHOLE: average cost of $\ln {\cal L}$ is
4096 Hz is X.  First iteration averages about 30 minutes in this configuration, second and subsequent iterations require
about 2 h.

>>>>

\begin{figure*}
% COMMAND
% plot_posterior_corner.py --posterior-file posterior_samples-4.dat --posterior-file posterior_samples-3.dat --posterior-file posterior_samples-2.dat --posterior-file posterior_samples-1.dat --parameter mc --parameter delta_mc --mc-range  '[1.19655,1.19682]' --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc_bns.xml.gz  --composite-file all.net --quantiles None --posterior-label 4 --posterior-label 3 --posterior-label 2 --posterior-label 1 --use-legend

\caption{\label{fig:BNS:MultiIterate}\textbf{Convergence of BNS analysis}: Results for marginal posterior distributions
  of our fiducial synthetic neutron star.  Solid contours show credible intervals; solid one-dimensional distributions
  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\lambda$ and
  value of the underlying marginalized likelihood evaluations.   The corresponding dotted curves show an analysis using
  only the $m=\pm 2$ modes \editremark{perform both} \emph{Left panel } Posterior distribution
  over  $\mc$ and
  $\delta=(m_1-m_2)/M$.    \emph{Right panel}: Marginal 1d CDFs of $\mc$, showing convergence.
\emph{Bottom left}: Marginal 1d empirical CDF of $\ln{\cal L}_{\rm marg}$, showing that at late times the proposed
candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$.
\emph{Bottom right panel}: $\ln {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
error budget early on, the Monte Carlo error is only shown for the last iteration.
}
\end{figure*}

\begin{figure*}
\caption{\label{fig:BNS:Comparison}\textbf{Comparison of GPU, CPU analysis}: \editremark{worthwhile}
}
\end{figure*}

% 
In our first test we compare two workflows, one with the original embarassingly-parallel RIFT code using $N_{\rm
  eval}=1$ and one with the
``batched'' GPU-accelerated code with $N_{\rm eval}=20$, analyzing data with $f_{\rm sample}=4096\unit{Hz}$.    
% Code configuration
In this example, we use $100$ initial points spread over the two mass dimensions $\mc_z,\delta$, adding $20$ evaluations per
iteration.   We
choose this simple configuration to facilitate visualization.     The initial coarse grid covers a region $\mc_z\in[1.1962,1.1970] M_\odot$ and $\delta \in [0,0.25]$,  insuring the posterior was smaller than our initial
coarse grid placement.     The top  panels of Figure  \ref{fig:BNS:MultiIterate} show posterior disributions derived
from the first several RIFT iterations, while the bottom panels show convergence diagnostics.
Figure \ref{fig:BNS:Comparison} compares results from the GPU and CPU codes on \editremark{identical test points: the
  initial grid}.\footnote{This comparison of $\ln {\cal L}_{\rm marg}$ is limited by Monte Carlo integration error and
  is provided  as a pedagogically convenient sanity check.  When comparing 
  $\ln {\cal L}_{\rm margT}$ results on identical data and parameters, the GPU and CPU implementations agree to
  significantly higher precision.}  The two results agree.
Because our analysis uses a post-Newtonian model to interpret a signal generated with SEOBNRv4, all results are offset
slightly from the synthetic signal's parameters

This specific workflow configuration  reduces overall core usage by maximizing GPU use per iteration: after the first
iteration, only one GPU is active.   
Specifically,  with the updated RIFT workflow used here, with one instance analyzing each 20 evaluations,  we use five
core+GPU pairs in the
first iteration, followed by one core+GPU for remaining iterations.
By contrast, with the original CPU-only RIFT code analyzing each $\lambda$ in parallel, this process requires roughly 20 core-minutes
per $\lambda$, using 100 cores in the first iteration and 20 cores in each
subsequent evalution.   
Note that  because   $N_{\rm eval}=20$ for the GPU is larger than the (hardware-and $f_{\rm sample}$ dependent)  speedup factor between the CPU and
GPU implementation, the overall wallclock time needed for a end-to-end
analysis analysis is larger for the GPU workflow.
We can achieve comparable or smaller turnaround time for an otherwise identical analysis by adjusting $N_{\rm eval}$ appropriate to the available hardware.

\editremark{modify output size: want more than 3000 samples, probably want 10k to make sure we are stable and to
  minimize MC error}
